---
title: "My First Parameterised Report"
author: "Mike K Smith"
date: "`r format(Sys.time(), '%d %B, %Y')`"
params: 
  endpoint: 
    value: HAMDTL17
    choices:
      - HAMDTL17
      - HAMATOTL
      - PGIIMP
output:
  html_document:
    df_print: paged
    toc: TRUE
    toc_float: TRUE
    toc_depth: 4
    code_download: TRUE
  pdf_document: default
  word_document: default
---
# Aims
This report shows a typical report where the tables, figures and numbers in the
report can be generated using R.

```{r loadTidyverse, warnings = FALSE, results="hide", message=FALSE, echo=FALSE}
library(tidyverse)
library(broom.mixed)
library(nlme)

## Hide code 
knitr::opts_chunk$set(echo = FALSE)
```

# Data Source
We're using a publically available dataset on depression. 
  
  * The data is from https://missingdata.lshtm.ac.uk/category/dia-working-group/example-data-sets/
(Website accessed 05 June 2018).  
  * The associated manuscript is https://www.ncbi.nlm.nih.gov/pubmed/15232330.  

```{r DataManipulation, message = FALSE, echo = FALSE, results = "hide", warning = FALSE}
## ----DataManipulation----------------------------------------------------
data <- haven::read_sas("../data/chapter15_example.sas7bdat")

data <- data %>%
  rename_all(funs(
    str_replace(string = ., pattern=params$endpoint, replacement="outcome")
  )) %>%
  bind_cols(data,.) %>%
  drop_na()


data %>%
  head(10)
```

# Outcomes
There are three outcomes in the dataset:   
  
  * HAM-D 17 item rating (`HAMDTL17`)  
  * Hamilton Anziety Rating Scale (`HAMATOTL`)  
  * Physician rating of global improvement (`PGIIMP`)  
  
This report summarises `r params$endpoint`. The parameter selected is passed
as a report parameter in the YAML header and can be over-ridden via  
`rmarkdown::render`.

# Exploratory data analysis
There are `r length(unique(data$PATIENT))` subjects in the dataset. Endpoints
are measured on four visits ranging from `r min(data$RELDAYS)` days to 
`r max(data$RELDAYS)` days. 

## Summary Statistics
The table below shows mean `r params$endpoint` by THERAPY and VISIT.
  
```{r data_summary}
data %>%
  group_by(THERAPY, VISIT) %>%
  summarise(n = n(),
            mean = mean(outcome),
            sd = sd(outcome),
            range = paste(
              min(outcome), 
              "-",
              max(outcome)
            ))
```
  
## Visualisation  
  
```{r graphical_EDA}
data <- data %>%
  mutate(WEEK = floor(RELDAYS/7))

ggplot(data = data, aes(x=jitter(WEEK), y=outcome, group=THERAPY, colour=THERAPY)) +
  geom_point(alpha=0.25) + 
  geom_smooth(method=lm) +
  labs(title=paste(params$endpoint,"by study week"),
       x="Week",
       y=params$endpoint) + 
  theme(legend.position = "bottom")
```
  
# Analysis
We fit a linear mixed effect model to the change in `r params$endpoint` over 
time using baseline as a covariate and look for difference in slope. We also 
allow for random between subject variability in the intercept and slope. The 
analysis examines three models (1/ no random effects, 2/ random intercept only,
3/ random intercept and random slope) and chooses the best based on "most 
significant". In a real analysis, other more appropriate model choice criteria 
should be used!

```{r linearModel}
data <- data %>%
  mutate(THERAPY = factor(THERAPY,
                          levels = c("PLACEBO","DRUG"))) 

fit <- lme(outcome ~ basval + RELDAYS, 
           data=data, 
           random = ~1 + RELDAYS | PATIENT,
           method = "ML",
           na.action = na.omit)

fit2 <- lme(outcome ~ basval + RELDAYS + THERAPY, 
           data=data, 
           random = ~1 | PATIENT,
           method = "ML",
           na.action = na.omit)

fit3 <- lme(outcome ~ basval + RELDAYS + THERAPY, 
           data=data, 
           random = ~1 + RELDAYS | PATIENT,
           method = "ML",
           na.action = na.omit)

modelComparison <- anova(fit, fit2, fit3)

pValue <- as.numeric(modelComparison$`p-value`)
signif <- pValue < 0.05
bestModel <- sum(signif, na.rm=TRUE)+1
bestFit <- eval(parse(text = glue::glue("fit{bestModel}")))
```

## Model Predictions
```{r model_predictions, out.width="100%"}
predData <- data %>%
            mutate(pred = predict(bestFit,level = 0),
                   ipred = predict(bestFit,level = 1))

broom.mixed::tidy(bestFit)

predData<- predData %>%
  group_by(THERAPY, WEEK) %>%
  summarise(MEAN = mean(outcome),
            pred = mean(pred),
            ipred = mean(ipred),
            STD = sd(outcome),
            N = n()) %>%
  filter( N > 1 ) %>%
  mutate(SE = (STD/sqrt(N)),
         LOWER = MEAN - 1.96*SE,
         UPPER = MEAN + 1.96*SE) 

ggplot(data = predData, aes(x = WEEK, y = MEAN, group=THERAPY, colour=THERAPY)) +
    facet_grid( ~ THERAPY) +
    geom_point() +
    geom_line() +
    geom_line(aes(x = WEEK, y = pred), colour = "red") + 
    geom_linerange(aes(x = WEEK, ymin = LOWER, ymax = UPPER)) +
    labs(title = glue::glue("Mean {params$endpoint} with CIs + mean population predictions")) +
    theme(legend.position = "bottom")

ggplot(data = predData, aes(x = WEEK, y = MEAN, group=THERAPY, colour=THERAPY)) +
    facet_grid( ~ THERAPY) +
    geom_point() +
    geom_line() +
    geom_line(aes(x = WEEK, y = ipred), colour = "red") + 
    geom_linerange(aes(x = WEEK, ymin = LOWER, ymax = UPPER)) +
    labs(title = glue::glue("Mean {params$endpoint} with CIs + mean individualised predictions")) +
    theme(legend.position = "bottom")
```

---

# Appendix 1 - Session information
```{r sessionInfo}
sessionInfo()
```
